### 2016-12-05 methodsPaper/lib/abstract.Rnw

Comparative risk assessment analyses to quantify the health potential
of active transportation, e.g., walking and cycling, have been
performed in the United States on a limited basis. Current
methodologies, which rely on extensive data collection and user
expertise, have estimated health benefits for select metropolitan
areas under a few alternative scenarios. These analyses have largely
been conducted using the Integrated Transport Health Impacts Model
(ITHIM) which was developed in the UK and has been adapted for use in
the US, primarily California. Here we present an implementation of the
active transportation component of ITHIM to model the health impacts
of modified active transportation levels. Using the newly developed
software package \package{} written in R, along with publicly
available national datasets, we quantify the change in disability
adjusted life years over a wide range of scenarios.  The R package
implemention of ITHIM's physical activity component provides a robust
environment for exploring the relationship between active
transportation and positive health outcomes.  A web-based
user-interface is available for users who want a first estimate for
the impact active transportation has on their
state/county/tract/region.

>>>>

Comparative risk assessments to quantify the health potential
of active transportation, e.g., walking and cycling, have been
performed in the United States on a limited basis. Current
methodologies, which rely on extensive data collection and user
expertise, have estimated health benefits for select metropolitan
areas under a few alternative scenarios. These analyses have largely
been conducted using the Integrated Transport Health Impacts Model
(ITHIM) which was developed in the UK and has been adapted for use in
the US, primarily California. Here we present an implementation of the
active transportation component of ITHIM to model the health impacts
of modified active transportation levels. Using the newly developed
software package \package{} written in R, along with publicly
available national datasets, we quantify the change in disability
adjusted life years over a wide range of scenarios.  The R package
implemention of ITHIM's physical activity component provides a robust
environment for exploring the relationship between active
transportation and positive health outcomes.  A web-based
user-interface is available for users who want a first estimate for
the impact active transportation has on their
state/region.

### 2016-12-05 methodsPaper/lib/introduction.Rnw

Regular exercise has been shown to provide health benefits in terms of
chronic disease prevention \cite{warburton2006}, cognitive function
\cite{hillman2008}, and overall well being and satisfaction with life
\cite{maher2013}. Still only half of Americans obtain the recommended
levels of aerobic activity, which is 2.5 hours per week. There is
considerable variability in the attainment of this goal by location
within the country. Supportive local built and policy environments are
important determinants of population-level activity
patterns. Particularly, as they facilitate and foster routine and
moderate physical activity, such as walking, local decisions on design
and policy can play a huge role in improving the nation's health.

@@

ITHIM uses a comparative risk assessment framework for the active
transport component of the model.  We improve on the existing
implementation of the active transport component by including the
distribution of non-travel-related activity, improving numerical
precision when computing the population attributable fraction and
creating a simple user-interface for the custom-built \R{} package
\package{}.

Here we present this implementation using publicly-available
datasets with coverage of the entire United States, and investigate
the sensitivity of the model outputs to variability in the input
parameters. The results of this work suggest that statistical modeling
and tools like the custom-built \R{} package \package{} can improve
the application of this useful methodology and increase its utility to
both practicioners and researchers.

>>>>

Regular exercise has been shown to provide health benefits in terms of
chronic disease prevention \cite{warburton2006}, cognitive function
\cite{hillman2008}, and overall well being and satisfaction with life
\cite{maher2013}. Still only half of Americans obtain the recommended
levels of aerobic activity of 2.5 hours per week. There is
considerable variability in the attainment of this goal by location
within the country. Supportive local built and policy environments are
important determinants of population-level activity
patterns. Particularly, as they facilitate and foster routine and
moderate physical activity, such as walking, local decisions on design
and policy can play a huge role in improving the nation's health.

@@

ITHIM uses a comparative risk assessment framework for the active
transport component of the model.  We improve on the existing
implementation of the active transport component by including the
distribution of non-travel-related activity, improving numerical
precision when computing the population attributable fraction and
creating a simple user-interface for the custom-built \R{} package
\package{}.  We present this implementation here by using it to
estimate the relationship between overall, nationwide disease burden
and active transportation time in the United States.

### 2016-12-05 methodsPaper/tex/materialsAndMethods.Rnw

As with the original ITHIM implementation we employ a comparative risk
analysis (CRA) across scenarios defined by mean active transportation
time, i.e., walking and cycling.  To do so we estimate the
proportional change in disease burden between a baseline model and an
alternative scenario, namely the population attributable fraction,
$\af$.  The population attributable fraction is defined below and
computed for each disease included in the model.

@@

Within each age-sex class the non-travel-related exposure is modeled
with a \logNormal{} distribution.  The age-sex specific mean ratios,
$r_{ij} = \frac{\mu_{ij}}{\mu_0}$ and $\gamma$ are estimated from the
American Time USe Survey data \cite{ATUS}.

>>>>

As with the original ITHIM implementation we, employed a comparative
risk assessment (CRA) across scenarios defined by mean active
transportation time, i.e., combined walking and cycling.  To do so we
investigate the proportional change in national disease burden between
the baseline model and a given alternative scenario, i.e., the
population attributable fraction, $\af$.  The population attributable
fraction is defined below and computed for each disease (ICD-10)
included in the model, namely Breast Cancer (C50), Colon Cancer
(C18-C21), Dementia (G30), Diabetes (E10-E14) and Cardiovascular
disease (I10, I12, I15, I20-I25, I30-I31, I33, I40).  We also include
the disability impacts of depression, though it is not listed as a
cause of death in CDC's data. For readability we omit an index
indicating disease.

@@

Within each age-sex class the non-travel-related exposure is modeled
with a \logNormal{} distribution.  The age-sex specific mean ratios,
$r_{ij} = \frac{\mu_{ij}}{\mu_0}$ and $\gamma$ are estimated from the
American Time Use Survey data \cite{ATUS}.
